### 2015-09-17 rscripts/location_clusters.Rmd
jp commit something

```{r,echo=F,include=F}
# http://stackoverflow.com/questions/18788748/fit-a-mixture-of-von-mises-distributions-in-r
psi <- peaks[measure == "usage", peak/24*pi] # usage angle
tht <- peaks[measure == "login", peak/24*pi] # login angle
phi <- peaks[measure == "logout", peak/24*pi] # logout angle
x0 <- cos(psi)
x1 <- sin(psi)*cos(tht)
x2 <- sin(psi)*sin(tht)*cos(phi)
x3 <- sin(psi)*sin(tht)*sin(phi)
pos <- cbind(x0,x1,x2,x3)
temporal_cluster <- data.table(location_id = peaks[measure == "usage"]$location_id, vMFcluster = predict(movMF(pos, 3, list(maxiter=10000, nruns=20)), pos), key="location_id")

@@

```{r, echo=F, fig.cap="Frequency Power Clusters"}
setkey(pwr, measure, location_id, variable)
peak_freq_cat <- pwr[,list(peak_freq = ifelse(variable[which.max(value)] == "1_perday","daily","multimode")), keyby=list(measure, location_id)]
tot_pwr <- pwr[,list(tot_pwr = sum(value)), keyby=list(measure, location_id)]
pwr_divisions <- tot_pwr[,{
  res <- as.list(quantile(tot_pwr, probs = c(1,2)/3))
  names(res) <- c("low","hi")
  res
}, keyby=measure]
pwr_cluster <- tot_pwr[pwr_divisions][peak_freq_cat][,
  list(peak_freq, pwr_cat = ifelse(tot_pwr < low,"low",ifelse(tot_pwr < hi, "med","hi"))),
  keyby=list(measure, location_id)
]
ggplot(pwr[pwr_cluster][,
  list(power=value, frequency=as.numeric(sub("_perday","", variable)), peak_freq, pwr_cat),
  by=list(location_id, measure)
]) + theme_bw() + aes(y = power, x=frequency, group=location_id, alpha=peak_freq, color = pwr_cat) + facet_grid(measure ~ .) + geom_line() + scale_x_continuous("per day frequency", breaks=c(1,2,3,4,6,8,12))
```

@@

@@

categorizer <- function(src.dt, cat, ref) {
  rle_expr <- paste0("rle(",cat,")")
  key_expr <- c("user_id",cat)
  mx_expr <- parse(text=paste0("list(main=",cat,"[which.max(prop)],share=max(prop))"))
  rle_dt <- src.dt[, rle(eval(parse(text=cat))), keyby=user_id]
  prop_ref <- src.dt[,list(prop=sum(logout-login)), keyby=key_expr][ref][,list(prop=prop/usage), keyby=key_expr]
  prop_lab <- prop_ref[, eval(mx_expr), by=user_id]
  opts <- sort(unique(rle_dt$values))
  names(opts) <- paste0("cat_", opts)
  fun <- categorizer_help(opts)
  ## cast data, maybe?
  digest <- rle_dt[, list(tot_visits = sum(lengths), mn_reps=mean(lengths), switches=.N), keyby=user_id][ref][prop_lab]
  list(rle_dt=rle_dt, prop_ref=prop_ref, prop_lab=prop_lab, digest=digest)
}

#lifetime_dts <- categorizer(cluster_rle_src, "lifetime_cat", ref)
#pwr_dts <- categorizer(cluster_rle_src, "pwr_clust", ref)
#vMF_dts <- categorizer(cluster_rle_src, "vMFcluster", ref)

# lifetime_rle <- cluster_rle_src[,rle(lifetime_cat),keyby=user_id]
# lifetime_prop <- cluster_rle_src[,list(prop=sum(logout-login)), keyby=list(user_id, lifetime_cat)][ref][,list(prop=prop/usage), keyby=list(user_id, lifetime_cat)]
# lifetime_cat_dt <- lifetime_prop[, list(main=lifetime_cat[which.max(prop)], majority=max(prop)>0.5), by=user_id]
# 
# pwr_rle <- cluster_rle_src[,rle(pwr_clust),keyby=user_id]
# pwr_prop <- cluster_rle_src[,list(prop=sum(logout-login)), keyby=list(user_id, pwr_clust)][ref][,list(prop=prop/usage), keyby=list(user_id, pwr_clust)]
# pwr_prop_dt <- pwr_prop[, list(main=pwr_clust[which.max(prop)], majority=max(prop)>0.5), by=user_id]
# 
# 
# vMF_rle <- cluster_rle_src[,rle(vMFcluster), keyby=user_id]
# vMF_prop <- cluster_rle_src[,list(prop=sum(logout-login)), keyby=list(user_id, vMFcluster)][ref][,list(prop=prop/usage), keyby=list(user_id, vMFcluster)]
# vMF_prop_dt <- vMF_prop[, list(main=vMFcluster[which.max(prop)], majority=max(prop)>0.5), by=user_id]
# 
# 
# digest_lifetime_rle <- lifetime_rle[,list(
#     tot_visits = sum(lengths),
#     lows=sum((values=="low")*lengths),
#     mids=sum((values=="mid")*lengths),
#     highs=sum((values=="high")*lengths),
#     mn_reps=mean(lengths),
#     switches = .N
#   ), keyby=user_id][ref][lifetime_cat_dt]
# digest_pwr_rle <- pwr_rle[,list(
#     tot_visits = sum(lengths),
#     lows=sum((values=="low")*lengths),
#     mids=sum((values=="mid")*lengths),
#     highs=sum((values=="high")*lengths),
#     mn_reps=mean(lengths),
#     switches = .N
#   ), keyby=user_id][ref][pwr_prop_dt]
# digest_vMF_rle <- vMF_rle[,list(
#     tot_visits = sum(lengths),
#     As=sum((values==1)*lengths),
#     Bs=sum((values==2)*lengths),
#     Cs=sum((values==3)*lengths),
#     mn_reps=mean(lengths),
#     switches = .N
#   ), keyby=user_id][ref][vMF_prop_dt]

@@

```{r, echo=F}
#ggplot(lifetime_dts$digest) + theme_bw() + aes(x=log10(switches/usage), fill=main, alpha=majority) + geom_bar() + labs(title="By Lifetime Categories") + scale_alpha_manual(values=c(0.6, 1))
#ggplot(pwr_dts$digest) + theme_bw() + aes(x=log10(switches/usage), fill=main, alpha=majority) + geom_bar() + labs(title="By Frequency Power Categories") + scale_alpha_manual(values=c(0.6, 1))
#ggplot(vMF_dts$digest) + theme_bw() + aes(x=log10(switches/usage), fill=factor(main), alpha=majority) + geom_bar() + labs(title="By von Mises-Fisher Categories") + scale_alpha_manual(values=c(0.6, 1))
```

>>>>

```{r,echo=F,include=F}
# http://stackoverflow.com/questions/18788748/fit-a-mixture-of-von-mises-distributions-in-r
psi <- peaks[measure == "usage", peak/24*pi] # usage angle
tht <- peaks[measure == "login", peak/24*pi] # login angle
phi <- peaks[measure == "logout", peak/24*pi] # logout angle
x0 <- cos(psi)
x1 <- sin(psi)*cos(tht)
x2 <- sin(psi)*sin(tht)*cos(phi)
x3 <- sin(psi)*sin(tht)*sin(phi)
pos <- cbind(x0,x1,x2,x3)

lls <- lapply(8:11, function(k) movMF(pos, k, list(maxiter=10000, nruns=20)))
bics <- lapply(lls, stats::BIC)

temporal_cluster <- data.table(location_id = peaks[measure == "usage"]$location_id, vMFcluster = predict(movMF(pos, 3, list(maxiter=10000, nruns=20)), pos), key="location_id")

alt_temporal_cluster <- data.table(location_id = peaks[measure == "usage"]$location_id, vMFcluster = predict(lls[[1]], pos), key="location_id")

@@

```{r, echo=F, fig.cap="Frequency Power Clusters"}
setkey(pwr, measure, location_id, variable)
peak_freq_cat <- pwr[,list(peak_freq = ifelse(variable[which.max(value)] == "1_perday","daily","multimode")), keyby=list(measure, location_id)]
tot_pwr <- pwr[,list(tot_pwr = sum(value)), keyby=list(measure, location_id)]
pwr_divisions <- tot_pwr[,{
  res <- as.list(quantile(tot_pwr, probs = c(1,2)/3))
  names(res) <- c("low","hi")
  res
}, keyby=measure]
pwr_cluster <- tot_pwr[pwr_divisions][peak_freq_cat][,
  list(peak_freq, pwr_cat = ifelse(tot_pwr < low,"lo",ifelse(tot_pwr < hi, "med","hi"))),
  keyby=list(measure, location_id)
]
ggplot(pwr[pwr_cluster][,
  list(power=value, frequency=as.numeric(sub("_perday","", variable)), peak_freq, pwr_cat),
  by=list(location_id, measure)
]) + theme_bw() + aes(y = power, x=frequency, group=location_id, alpha=peak_freq, color = pwr_cat) + facet_grid(measure ~ .) + geom_line() + scale_x_continuous("per day frequency", breaks=c(1,2,3,4,6,8,12))
```

@@

location diversity - share of locations in each type, in each dimension.  We know by construction that location diversity is uniform for low, medium, and high usage per lifetime categorization: we created the categories based on quantiles.  Same is true for use variation frequency power.  For the clusters of peak utilization characteristics, however,

```{r, echo=F}
multinomtest <- function(xs, nh=rep(1/length(xs), times=length(xs))) {
  N <- sum(xs)
  2*N*log(N) - 2*sum(xs*log(nh/xs))
}
agg.dt[,.N,by=lifetime_cat][,list(lifetime_cat, share=N/sum(N))]
agg.dt[,.N,by=pwr_clust][,list(pwr_clust, share=N/sum(N))]
agg.dt[,.N,by=vMFcluster][,list(vMFcluster, share=N/sum(N))]
```

@@

categorizer <- function(src.dt, cat, ref) {
  rle_expr <- paste0("rle(",cat,")")
  key_expr <- c("user_id",cat)
  mx_expr <- parse(text=paste0("list(main=",cat,"[which.max(prop)],share=max(prop))"))
  rle_dt <- src.dt[, rle(eval(parse(text=cat))), keyby=user_id]
  prop_ref <- src.dt[,list(prop=sum(logout-login)), keyby=key_expr][ref][,list(prop=prop/usage), keyby=key_expr]
  prop_lab <- prop_ref[, eval(mx_expr), by=user_id]
  chunks <- data.table(dcast(rle_dt, user_id ~ values, value.var="lengths", fun.aggregate = sum), key="user_id")
  ## cast data, maybe?
  digest <- rle_dt[, list(tot_visits = sum(lengths), mn_reps=mean(lengths), switches=.N), keyby=user_id][ref][prop_lab][chunks]
  list(rle_dt=rle_dt, prop_ref=prop_ref, prop_lab=prop_lab, digest=digest)
}

lifetime_dts <- categorizer(cluster_rle_src, "lifetime_cat", ref)
pwr_dts <- categorizer(cluster_rle_src, "pwr_clust", ref)
vMF_dts <- categorizer(cluster_rle_src, "vMFcluster", ref)

@@

\newpage\blandscape
```{r, echo=F, fig.width=11.25, fig.height=7, fig.cap="distribution of main location type usage"}
digests <- rbind(
  lifetime_dts$digest[,list(switch_rate = switches/usage, main, share, type="usage")],
  pwr_dts$digest[,list(switch_rate = switches/usage, main, share, type="power")],
  vMF_dts$digest[,list(switch_rate = switches/usage, main, share, type="peaktime")]
)
ggplot(digests) + facet_grid(type ~ .) + theme_bw() + aes(x=log10(switch_rate), fill=main, alpha=factor(floor(share*10))) + geom_bar(binwidth=0.1) + scale_alpha_manual(breaks=0:10, values=seq(0.5,1,length.out = 11))
ggplot(digests) + facet_grid(type ~ .) + theme_bw() + aes(x=log10(switch_rate), fill=main, alpha=factor(floor(share*10))) + geom_bar(binwidth=0.1, position="fill") + scale_alpha_manual(breaks=0:10, values=seq(0.5,1,length.out = 11))
```
\elandscape\newpage

\newpage\blandscape
```{r, echo=F, fig.width=11.25, fig.height=7, fig.cap="user categorization, by primary location type utilization"}
plot_ref <- ref[lifetime_dts$digest[, list(lifetime_main = main), keyby=user_id]][pwr_dts$digest[, list(pwr_main = main), keyby=user_id]][vMF_dts$digest[, list(peak_main = main), keyby=user_id]]
#thing[, list(user_id, mains=interaction(lifetime_main, pwr_main, peak_main))][,.N,by=mains]
ggplot(plot_ref[life > 365*24*60*60 & ul > 2 & ul < 20, list(y=log10(usage), x=log10(life), pwr_main, lifetime_main, peak_main)]) +facet_grid(pwr_main ~ lifetime_main) + aes(x=x,y=y, color=factor(peak_main)) + geom_point(alpha=0.3) + theme_bw()
```
\elandscape\newpage
